---
title: "Download and plot SST"
format: html
editor: visual
---

### Load packages

```{r}
library(httr) #for downloads
library(here) #for project folder structure
library(tidyverse)
library(pbapply) #for progress bars
library(raster) #to load .nc files
library(lubridate) #fast date change function
library(sf) # mapping
library(ggspatial) # mapping
library(viridis) #plots
library(patchwork) #plots
library(ncdf4) # read files
library(patchwork)
```

### Load my functions

```{r}
source(here("scripts/Andi_DHW_functions.R"))
```

### Climatology

Download global climatology data (MMM). Un-comment this code to download the data:

```{r}
# url_climatology = "https://www.star.nesdis.noaa.gov/pub/sod/mecb/crw/data/5km/v3.1_op/climatology/nc/ct5km_climatology_v3.1.nc"
# 
# httr::GET(url_climatology,
#             httr::write_disk(here("data/climatology/ct5km_climatology_v3.1.nc"),
#                              overwrite = T),
#             httr::add_headers('Accept-Encoding' = 'gzip, deflate'),
#             httr::progress("down"),
#             httr::timeout(600))
```

Read the downloaded data

```{r}
climatology <- raster::brick(here("data/climatology/ct5km_climatology_v3.1.nc"), varname = "sst_clim_mmm")
```

### Download data

I already downloaded the data, but just in case you need to do it again:

The `download_all()` function downloads the data in batches of 3 years (as default value, other values are also possible)

The data will be downloaded to the folder you specify

You can either download the SST data for an area, then use the range of `lat1`/`lat2` and `lon1`/`lon2` values. If you just want one location, just enter the same values in `lat1`/`lat2` and `lon1`/`lon2` .

One `.nc` file will be downloaded per 3 years.

### Moorea

Get box around Moorea of 15 km

```{r}
moorea_lat = -17.536273
moorea_lon = -149.832730

moorea_box = bounding_box(moorea_lat, moorea_lon, 15, 15)

```

Visualise

```{r}
map_FrenchPolynesia <- st_read(dsn = here("data/maps/polynesia_map/PYF_adm0.shp")) %>%
  st_as_sf()
```

This area will be downloaded\

```{r}
map_moorea <- ggplot() +
  geom_sf(data = map_FrenchPolynesia,
          fill = "antiquewhite",
          col = "black")+
  
  coord_sf(xlim = c(moorea_box$lon_min, moorea_box$lon_max),
           ylim = c(moorea_box$lat_min, moorea_box$lat_max))+
  labs(x=NULL, y = NULL, title = "Moorea")+
  theme_minimal()

map_moorea
```

```{r}
# download_all(lat1 = moorea_box$lat_min,
#              lat2 = moorea_box$lat_max,
#              lon1 = moorea_box$lon_min,
#              lon2 = moorea_box$lon_max,
#              day1 = "1985-04-01",#min date
#              day2 = "2024-05-22",
#              folder = here("data/ncdf/moorea_sst"))
```

Read

```{r}
dat_moorea_sst <- read_nc(here("data/ncdf/moorea_sst"),
                      varname = "CRW_SST", 
                      values_to = "sst")
```

Similarly, you can download DHW:

```{r}
# download_all(lat1 = moorea_box$lat_min,
#              lat2 = moorea_box$lat_max,
#              lon1 = moorea_box$lon_min,
#              lon2 = moorea_box$lon_max,
#              day1 = "1985-04-01",
#              day2 = "2024-05-22",
#              parameter = "CRW_DHW",
#              folder = here("data/ncdf/moorea_dhw"))
```

Read

```{r}
dat_moorea_dhw <- read_nc(here("data/ncdf/moorea_dhw"),
                      varname = "CRW_DHW", 
                      values_to = "dhw")
```

Summarise over area and combine data

```{r}
dat_moorea_sst <- dat_moorea_sst %>% 
  group_by(date) %>% 
  summarise(sst = mean(sst, na.rm = T))

dat_moorea_dhw <- dat_moorea_dhw %>% 
  group_by(date) %>% 
  summarise(dhw = mean(dhw, na.rm = T))

dat_moorea <- left_join(dat_moorea_sst, dat_moorea_dhw, by = "date")

```

Get MMM

```{r}
mmm_moorea <- subset_ncdf(climatology,
                          lat_min = moorea_box$lat_min,
                          lat_max = moorea_box$lat_max,
                          lon_min = moorea_box$lon_min,
                          lon_max = moorea_box$lon_max)

# get mean
mmm_moorea <- mean(mmm_moorea$mmm, na.rm = T) +1
  
```

```{r}
plot_sst_moorea <- dat_moorea %>% 
  ggplot(aes(x = date, y = sst))+
  geom_hline(yintercept = mmm_moorea, linetype = "dashed", col = "#D86D62")+
  geom_line()+
  annotate(geom = "text",
           x = as.Date(-Inf),
           y = mmm_moorea, 
           label = paste(round(mmm_moorea,1),"°C"), 
           col = "#D86D62", vjust = -.3, hjust = -.5)+
  scale_x_date(date_breaks = "4 years", date_labels =  "%Y") +
  labs(x = NULL, y = "SST (°C)", title = "A) Moorea")+
  theme_classic()


plot_sst_moorea
```

DHW

```{r}
#summarise data
dat_moorea_dhwS <- dat_moorea %>% 
  mutate(year = factor(year(date))) %>% 
  group_by(year) %>% 
  summarise(max_dhw = max(dhw, na.rm = T)) %>% 
  filter(max_dhw > 1)  # I think that's what they did
```

```{r}
# # as a test, plot DHW in 2016
# 
# dat_moorea_maxdhw<- dat_moorea_dhw %>% 
#                   mutate(year = year(date)) %>% 
#                   group_by(year, lat, lon) %>% 
#                   summarise(dhw = max(dhw, na.rm = T))
# 
# ggplot() +
#   geom_raster(data = dat_moorea_maxdhw %>% 
#                 filter(year == 2024),
#               aes(y = lat, x = lon, fill = dhw)) +
#   geom_sf(data = map_FrenchPolynesia,
#           fill = "antiquewhite",
#           col = "black")+
#   coord_sf(xlim = c(moorea_box$lon_min, moorea_box$lon_max),
#            ylim = c(moorea_box$lat_min, moorea_box$lat_max))+
#   labs(x=NULL, y = NULL, title = "DHW Moorea 2016")+
#   scale_fill_viridis()+
#   theme_minimal()


```

```{r}
#plot
plot_dhw_moorea <- dat_moorea_dhwS %>% 
  ggplot(aes(x = year, y = max_dhw))+
  geom_bar(stat = "identity", fill = "#D86D62")+
  geom_text(aes(x = year,
               y = max_dhw,
               label = round(max_dhw, 1)),
            vjust = -.3)+
  labs(x = "Bleaching events", y = "DHW (°C weeks)")+
 theme_classic()


plot_dhw_moorea
```

```{r}
plot_moorea <- plot_sst_moorea + plot_dhw_moorea +
  plot_layout(widths = c(2, 1))
plot_moorea
```

### Fakarava

Get box around Fakarava of 15 km

```{r}
fakarava_lat = -16.3132050
fakarava_lon = -145.621565

fakarava_box = bounding_box(fakarava_lat, fakarava_lon, 15, 15)

```

Visualise This area will be downloaded\

```{r}
map_fakarava <- ggplot() +
  geom_sf(data = map_FrenchPolynesia,
          fill = "antiquewhite",
          col = "black")+
  
  coord_sf(xlim = c(fakarava_box$lon_min, fakarava_box$lon_max),
           ylim = c(fakarava_box$lat_min, fakarava_box$lat_max))+
  labs(x=NULL, y = NULL, title = "Fakarava")+
  theme_minimal()

map_fakarava
```

```{r}
# download_all(lat1 = fakarava_box$lat_min,
#              lat2 = fakarava_box$lat_max,
#              lon1 = fakarava_box$lon_min,
#              lon2 = fakarava_box$lon_max,
#              day1 = "1985-04-01",#min date
#              day2 = "2024-05-22",
#              folder = here("data/ncdf/fakarava_sst"))
```

Read

```{r}
dat_fakarava_sst <- read_nc(here("data/ncdf/fakarava_sst"),
                      varname = "CRW_SST", 
                      values_to = "sst")
```

Similarly, you can download DHW:

```{r}
# download_all(lat1 = fakarava_box$lat_min,
#              lat2 = fakarava_box$lat_max,
#              lon1 = fakarava_box$lon_min,
#              lon2 = fakarava_box$lon_max,
#              day1 = "1985-04-01",
#              day2 = "2024-05-22",
#              parameter = "CRW_DHW",
#              folder = here("data/ncdf/fakarava_dhw"))
```

Read

```{r}
dat_fakarava_dhw <- read_nc(here("data/ncdf/fakarava_dhw"),
                      varname = "CRW_DHW", 
                      values_to = "dhw")
```

Summarise over area and combine data

```{r}
dat_fakarava_sst <- dat_fakarava_sst %>% 
  group_by(date) %>% 
  summarise(sst = mean(sst, na.rm = T))

dat_fakarava_dhw <- dat_fakarava_dhw %>% 
  group_by(date) %>% 
  summarise(dhw = mean(dhw, na.rm = T))

dat_fakarava <- left_join(dat_fakarava_sst, dat_fakarava_dhw, by = "date")

```

Get MMM

```{r}
mmm_fakarava <- subset_ncdf(climatology,
                          lat_min = fakarava_box$lat_min,
                          lat_max = fakarava_box$lat_max,
                          lon_min = fakarava_box$lon_min,
                          lon_max = fakarava_box$lon_max)

# get mean
mmm_fakarava <- mean(mmm_fakarava$mmm, na.rm = T) +1
  
```

```{r}
plot_sst_fakarava <- dat_fakarava %>% 
  ggplot(aes(x = date, y = sst))+
  geom_hline(yintercept = mmm_fakarava, linetype = "dashed", col = "#D86D62")+
  geom_line()+
  annotate(geom = "text",
           x = as.Date(-Inf),
           y = mmm_fakarava, 
           label = paste(round(mmm_fakarava,1),"°C"), 
           col = "#D86D62", vjust = -.3, hjust = -.5)+
  scale_x_date(date_breaks = "4 years", date_labels =  "%Y") +
  labs(x = NULL, y = "SST (°C)", title = "B) Fakarava")+
  theme_classic()


plot_sst_fakarava
```

DHW

```{r}
#summarise data
dat_fakarava_dhwS <- dat_fakarava %>% 
  mutate(year = factor(year(date))) %>% 
  group_by(year) %>% 
  summarise(max_dhw = max(dhw, na.rm = T)) %>% 
  filter(max_dhw > 1)  # I think that's what they did
  

#plot
plot_dhw_fakarava <- dat_fakarava_dhwS %>% 
  ggplot(aes(x = year, y = max_dhw))+
  geom_bar(stat = "identity", fill = "#D86D62")+
  geom_text(aes(x = year,
               y = max_dhw,
               label = round(max_dhw, 1)),
            vjust = -.3)+
  labs(x = "Bleaching events", y = "DHW (°C weeks)")+
 theme_classic()


plot_dhw_fakarava
```

```{r}
plot_fakarava <- plot_sst_fakarava + plot_dhw_fakarava +
  plot_layout(widths = c(2, 1))
plot_fakarava
```

### Tikehau

Get box around Tikehau of 15 km

```{r}
tikehau_lat = -14.9966195
tikehau_lon = -148.177199

tikehau_box = bounding_box(tikehau_lat, tikehau_lon, 15, 15)

```

Visualise This area will be downloaded\

```{r}
map_tikehau <- ggplot() +
  geom_sf(data = map_FrenchPolynesia,
          fill = "antiquewhite",
          col = "black")+
  
  coord_sf(xlim = c(tikehau_box$lon_min, tikehau_box$lon_max),
           ylim = c(tikehau_box$lat_min, tikehau_box$lat_max))+
  labs(x=NULL, y = NULL, title = "Tikehau")+
  theme_minimal()

map_tikehau
```

```{r}
# download_all(lat1 = tikehau_box$lat_min,
#              lat2 = tikehau_box$lat_max,
#              lon1 = tikehau_box$lon_min,
#              lon2 = tikehau_box$lon_max,
#              day1 = "1985-04-01",#min date
#              day2 = "2024-05-22",
#              folder = here("data/ncdf/tikehau_sst"))
```

Read

```{r}
dat_tikehau_sst <- read_nc(here("data/ncdf/tikehau_sst"),
                      varname = "CRW_SST", 
                      values_to = "sst")
```

Similarly, you can download DHW:

```{r}
# download_all(lat1 = tikehau_box$lat_min,
#              lat2 = tikehau_box$lat_max,
#              lon1 = tikehau_box$lon_min,
#              lon2 = tikehau_box$lon_max,
#              day1 = "1985-04-01",
#              day2 = "2024-05-22",
#              parameter = "CRW_DHW",
#              folder = here("data/ncdf/tikehau_dhw"))
```

Read

```{r}
dat_tikehau_dhw <- read_nc(here("data/ncdf/tikehau_dhw"),
                      varname = "CRW_DHW", 
                      values_to = "dhw")
```

Summarise over area and combine data

```{r}
dat_tikehau_sst <- dat_tikehau_sst %>% 
  group_by(date) %>% 
  summarise(sst = mean(sst, na.rm = T))

dat_tikehau_dhw <- dat_tikehau_dhw %>% 
  group_by(date) %>% 
  summarise(dhw = mean(dhw, na.rm = T))

dat_tikehau <- left_join(dat_tikehau_sst, dat_tikehau_dhw, by = "date")

```

Get MMM

```{r}
mmm_tikehau <- subset_ncdf(climatology,
                          lat_min = tikehau_box$lat_min,
                          lat_max = tikehau_box$lat_max,
                          lon_min = tikehau_box$lon_min,
                          lon_max = tikehau_box$lon_max)

# get mean
mmm_tikehau <- mean(mmm_tikehau$mmm, na.rm = T) +1
  
```

```{r}
plot_sst_tikehau <- dat_tikehau %>% 
  ggplot(aes(x = date, y = sst))+
  geom_hline(yintercept = mmm_tikehau, linetype = "dashed", col = "#D86D62")+
  geom_line()+
  annotate(geom = "text",
           x = as.Date(-Inf),
           y = mmm_tikehau, 
           label = paste(round(mmm_tikehau,1),"°C"), 
           col = "#D86D62", vjust = -.3, hjust = -.5)+
  scale_x_date(date_breaks = "4 years", date_labels =  "%Y") +
  labs(x = NULL, y = "SST (°C)", title = "C) Tikehau")+
  theme_classic()


plot_sst_tikehau
```

DHW

```{r}
#summarise data
dat_tikehau_dhwS <- dat_tikehau %>% 
  mutate(year = factor(year(date))) %>% 
  group_by(year) %>% 
  summarise(max_dhw = max(dhw, na.rm = T)) %>% 
  filter(max_dhw > 1)  # I think that's what they did
  

#plot
plot_dhw_tikehau<- dat_tikehau_dhwS %>% 
  ggplot(aes(x = year, y = max_dhw))+
  geom_bar(stat = "identity", fill = "#D86D62")+
  geom_text(aes(x = year,
               y = max_dhw,
               label = round(max_dhw, 1)),
            vjust = -.3)+
  labs(x = "Bleaching events", y = "DHW (°C weeks)")+
 theme_classic()


plot_dhw_tikehau
```

```{r}
plot_tikehau <- plot_sst_tikehau + plot_dhw_tikehau +
  plot_layout(widths = c(2, 1))

plot_tikehau
```

### Bora Bora

Get box around Bora Bora of 15 km

```{r}
bora_lat = -16.4920754
bora_lon = -151.740538

bora_box = bounding_box(bora_lat, bora_lon, 15, 15)

```

Visualise This area will be downloaded\

```{r}
map_bora <- ggplot() +
  geom_sf(data = map_FrenchPolynesia,
          fill = "antiquewhite",
          col = "black")+
  
  coord_sf(xlim = c(bora_box$lon_min, bora_box$lon_max),
           ylim = c(bora_box$lat_min, bora_box$lat_max))+
  labs(x=NULL, y = NULL, title = "Bora Bora")+
  theme_minimal()

map_bora
```

```{r}
# download_all(lat1 = bora_box$lat_min,
#              lat2 = bora_box$lat_max,
#              lon1 = bora_box$lon_min,
#              lon2 = bora_box$lon_max,
#              day1 = "1985-04-01",#min date
#              day2 = "2024-05-22",
#              folder = here("data/ncdf/bora_sst"))
```

Read

```{r}
dat_bora_sst <- read_nc(here("data/ncdf/bora_sst"),
                      varname = "CRW_SST", 
                      values_to = "sst")
```

Similarly, you can download DHW:

```{r}
# download_all(lat1 = bora_box$lat_min,
#              lat2 = bora_box$lat_max,
#              lon1 = bora_box$lon_min,
#              lon2 = bora_box$lon_max,
#              day1 = "1985-04-01",
#              day2 = "2024-05-22",
#              parameter = "CRW_DHW",
#              folder = here("data/ncdf/bora_dhw"))
```

Read

```{r}
dat_bora_dhw <- read_nc(here("data/ncdf/bora_dhw"),
                      varname = "CRW_DHW", 
                      values_to = "dhw")
```

Summarise over area and combine data

```{r}
dat_bora_sst <- dat_bora_sst %>% 
  group_by(date) %>% 
  summarise(sst = mean(sst, na.rm = T))

dat_bora_dhw <- dat_bora_dhw %>% 
  group_by(date) %>% 
  summarise(dhw = mean(dhw, na.rm = T))

dat_bora <- left_join(dat_bora_sst, dat_bora_dhw, by = "date")

```

Get MMM

```{r}
mmm_bora <- subset_ncdf(climatology,
                          lat_min = bora_box$lat_min,
                          lat_max = bora_box$lat_max,
                          lon_min = bora_box$lon_min,
                          lon_max = bora_box$lon_max)

# get mean
mmm_bora <- mean(mmm_bora$mmm, na.rm = T) +1
  
```

```{r}
plot_sst_bora <- dat_bora %>% 
  ggplot(aes(x = date, y = sst))+
  geom_hline(yintercept = mmm_bora, linetype = "dashed", col = "#D86D62")+
  geom_line()+
  annotate(geom = "text",
           x = as.Date(-Inf),
           y = mmm_bora, 
           label = paste(round(mmm_bora,1),"°C"), 
           col = "#D86D62", vjust = -.3, hjust = -.5)+
  scale_x_date(date_breaks = "4 years", date_labels =  "%Y") +
  labs(x = NULL, y = "SST (°C)", title = "D) Bora Bora")+
  theme_classic()


plot_sst_bora
```

DHW

```{r}
#summarise data
dat_bora_dhwS <- dat_bora %>% 
  mutate(year = factor(year(date))) %>% 
  group_by(year) %>% 
  summarise(max_dhw = max(dhw, na.rm = T)) %>% 
  filter(max_dhw > 1)  # I think that's what they did
  

#plot
plot_dhw_bora<- dat_bora_dhwS %>% 
  ggplot(aes(x = year, y = max_dhw))+
  geom_bar(stat = "identity", fill = "#D86D62")+
  geom_text(aes(x = year,
               y = max_dhw,
               label = round(max_dhw, 1)),
            vjust = -.3)+
  labs(x = "Bleaching events", y = "DHW (°C weeks)")+
 theme_classic()


plot_dhw_bora
```

```{r}
plot_bora <- plot_sst_bora + plot_dhw_bora +
  plot_layout(widths = c(2, 1))
plot_bora
```

### combine plots

```{r}
plots_all <- plot_moorea/plot_fakarava/plot_tikehau/plot_bora
plots_all
```

```{r}
ggsave(filename = "plot_sst_dhw.pdf", 
       plot = plots_all, 
       path = here("plots"),
       width = 30, height = 40, units = "cm")
```
